\section*{Problem 5: Knapsack with Partition Constraints}

Consider the following variant of the Knapsack problem. The input consists of $n$ items $I = \{1, \ldots, n\}$, where item $i \in I$ has a weight $w_i \in \mathbb{Z}^+$ and a value $v_i \in \mathbb{Z}^+$. The items are partitioned into disjoint sets $S_1, S_2, \ldots, S_m$; that is, $\bigcup_{j \in [m]} S_j = I$ and $S_j \cap S_\ell = \emptyset$ for all $j \neq \ell \in [m]$. Moreover, each set $S_j$ has a bound $k(j)$. Also, we are given a knapsack capacity $B \in \mathbb{Z}^+$. A solution for the problem is a subset of items $P \subseteq I$ such that $\sum_{i \in P} w_i \leq B$ and $|P \cap S_j| \leq k(j)$ for all $j \in [m]$. The value of the solution $P$ is $\sum_{i \in P} v_i$. The objective is to find a solution $P$ of maximal value. Obtain an FPTAS for the problem.

\subsection*{Overview}
We will maintain a \emph{domination list} (a list of nondominated $(\text{weight},\text{value})$ pairs).
A pair $(w,v)$ \emph{dominates} $(w',v')$ if $w \le w'$ and $v \ge v'$.
Dominated pairs can be discarded without affecting optimality.

For the FPTAS, we will apply the standard scaling technique (as in the usual knapsack FPTAS) and run the same domination-list DP on the scaled instance.

\subsection*{Algorithm Description}
\paragraph{Domination-list primitives.}
\begin{itemize}
    \item $\textsc{Shift}(L, \Delta w, \Delta v) = \{(w+\Delta w, v+\Delta v) : (w,v)\in L\}$.
    \item $\textsc{Prune}(L)$ removes all dominated pairs from $L$ (and may also remove pairs with $w>B$). One implementation: sort pairs by increasing weight; sweep, keeping a pair only if its value is strictly larger than the maximum value seen so far. Its complexity is $O(|L| \log |L|)$ due to sorting.
    \item $\textsc{Convolve}(L_1, L_2)$ computes all pairwise sums of pairs from $L_1$ and $L_2$: \[ \textsc{Convolve}(L_1, L_2) = \{(w_1 + w_2, v_1 + v_2) : (w_1, v_1) \in L_1, (w_2, v_2) \in L_2\} \]. Its complexity is $O(|L_1| \cdot |L_2|)$.
\end{itemize}

\begin{algorithm}[H]
\caption{Knapsack with Partition Constraints (Domination Lists)}
\begin{algorithmic}[1]
\Function{Knapsack-Partition}{$I, S_1, S_2, \ldots, S_m, k, B, \text{val}(\cdot)$}
    \State $L \gets \{(0,0)\}$ \Comment{global nondominated set of feasible (weight,value) pairs}
    \For{each partition $S_j$}
        \State \Comment{Compute nondominated options using items from $S_j$ with at most $k(j)$ picks}
        \State $L^{(0)} \gets \{(0,0)\}$
        \For{$c=1$ to $k(j)$} $L^{(c)} \gets \emptyset$ \EndFor
        \For{each item $i \in S_j$}
            \For{$c$ from $k(j)$ down to $1$}
                \State $L^{(c)} \gets \textsc{Prune}\Big( L^{(c)} \cup \textsc{Shift}(L^{(c-1)}, w_i, \text{val}(i))\Big)$
            \EndFor
        \EndFor
        \State \Comment{Combine the chosen items from $S_j$ with the global list}
        \State $L_{\text{next}} \gets \emptyset$
        \For{$c=0$ to $k(j)$}
            \State $L_{\text{next}} \gets \textsc{Prune}(L_{\text{next}} \cup \textsc{Convolve}(L, L^{(c)}))$
        \EndFor
        \State $L \gets L_{\text{next}}$
    \EndFor
    \State \Return $\max\{v : (w,v)\in L\}$
\EndFunction
\end{algorithmic}
\end{algorithm}


\subsection*{Proof of Correctness}
Without loss of generality, assume $\forall j,\ k(j) \le |S_j|$ (otherwise replace $k(j)$ by $|S_j|$).

\begin{lemma}
    After processing partitions $S_1,\dots,S_j$, the list $L$ contains exactly the nondominated pairs $(w,v)$ such that there exists a feasible selection of items from the first $j$ partitions of total weight $w$ and total value $v$ (respecting all bounds).
\end{lemma}

\begin{proof}
    \textbf{Base case ($j=0$):} $L=\{(0,0)\}$ is the unique feasible pair (select nothing), and it is nondominated.

    \textbf{Inductive step:} Assume the claim holds after $S_1,\dots,S_{j-1}$. The lists $L^{(c)}$ are constructed with an explicit counter $c$, and $\textsc{Prune}$ only removes dominated pairs, hence each $L^{(c)}$ represents exactly the nondominated achievable pairs using exactly $c$ items from $S_j$. 
    When convolving $L^{(c)}$ with the current global list $L$, we consider all ways to add $c$ items from $S_j$ to previously selected items (from partitions $1$ to $j-1$). The final pruning step ensures that only nondominated pairs remain in $L_{\text{next}}$. Thus, after processing $S_j$, $L$ contains exactly the nondominated pairs corresponding to feasible selections from the first $j$ partitions.
\end{proof}

\begin{theorem}
    The algorithm returns the optimal objective value for the knapsack with partition constraints problem.
\end{theorem}

\begin{proof}
    After processing all partitions, every feasible solution corresponds to some pair in the (unpruned) constructed set, and pruning preserves at least one representative of every attainable optimum value (for some weight). Thus $\max\{v:(w,v)\in L\}$ equals the optimal objective value under capacity $B$ and the partition bounds.
\end{proof}

\subsection*{Time Complexity}
Let $k_{\max}=\max_{j\in[m]} k(j)\le n$. Throughout the algorithm, we discard pairs with $w>B$.
Crucially, after a $\textsc{Prune}$ step the list contains only nondominated pairs, hence if we sort the remaining pairs by nondecreasing weight, their values are \emph{strictly increasing}. Therefore, for \emph{integer} weights and values, every pruned list has size at most
\[
M \;:=\; \min\{B+1,\,V_{\text{sum}}+1\},
\]
where $V_{\text{sum}}$ is an upper bound on the maximum attainable value in that run (e.g., $V_{\text{sum}}=\sum_{i\in I} v_i$ in the exact DP, or $V_{\text{sum}}=\sum_{i\in I} v'_i$ in the scaled DP).

In one partition $S_j$:
\begin{itemize}
    \item Updating the $L^{(c)}$ lists: for each item we perform at most $k(j)\le k_{\max}$ unions/shifts followed by $\textsc{Prune}$. Each $\textsc{Prune}$ costs $O(M\log M)$, so this part costs $O(|S_j|\cdot k(j)\cdot M\log M)$.
    \item Combining into $L_{\text{next}}$: we perform at most $k(j)+1$ convolutions, each costing $O(M^2)$, plus a prune costing $O(M\log M)$. Hence this part costs $O(k(j)\cdot M^2)$.
\end{itemize}
Summing over all partitions gives total running time
\[
O\Big(n\cdot k_{\max}\cdot M\log M \;+\; m\cdot k_{\max}\cdot M^2\Big),
\]
which is pseudo-polynomial (polynomial in $n,m,k_{\max}$ and in $M$).

\subsection*{FPTAS Construction}
To obtain an FPTAS, we apply the standard value-scaling technique and run the above DP on the scaled instance.

Let $p(i)\in[m]$ be the unique partition index such that $i\in S_{p(i)}$. Define
\[
V_{\max}^{\text{feas}} \;:=\; \max\{v_i : w_i\le B \text{ and } k(p(i))\ge 1\}.
\]
If no such item exists, then every feasible solution has value $0$, and we can return $0$.

Fix $\epsilon\in(0,1)$ and set
\[
K \;:=\; \frac{\epsilon\cdot V_{\max}^{\text{feas}}}{n},
\qquad
v'_i \;:=\; \Big\lfloor \frac{v_i}{K}\Big\rfloor \;\;\; \text{for all } i\in I.
\]

\begin{algorithm}[H]
\caption{FPTAS for Knapsack with Partition Constraints}
    \begin{algorithmic}[1]
    \Function{FPTAS-Knapsack-Partition}{$I, S_1, S_2, \ldots, S_m, k, B$}
        \State Scale values: for each $i \in I$, set $v'_i = \lfloor v_i / K \rfloor$
        \State Compute a solution $\widehat{P}$ that maximizes $\sum_{i\in P} v'_i$ among feasible sets $P$
        \State \hspace{1em}(by running \Call{Knapsack-Partition}{$I,S_1,\ldots,S_m,k,B,\text{val}(\cdot)$} with $\text{val}(i)=v'_i$ and keeping predecessor pointers)
        \State \Return $\sum_{i\in \widehat{P}} v_i$
    \EndFunction
    \end{algorithmic}
\end{algorithm}

\begin{proof}
    \textbf{Running time:} In the scaled instance, $V'_{\text{sum}}:=\sum_{i\in I} v'_i$ satisfies
    \[
    V'_{\text{sum}}
    \;\le\;
    \sum_{i\in I}\frac{v_i}{K}
    \;\le\;
    \frac{n\cdot V_{\max}^{\text{feas}}}{K}
    \;=\;
    \frac{n^2}{\epsilon}.
    \]
    Hence in the scaled DP we have $M\le \min\{B+1, V'_{\text{sum}}+1\}\le \frac{n^2}{\epsilon}+1$, and by the complexity bound proved above the running time is polynomial in $n$, $m$, and $1/\epsilon$.
\end{proof}

\begin{proof}
    \textbf{Approximation Ratio:} For the optimal set $P^*$, the difference between the true value and the scaled value is:
    \[
    \sum_{i \in P^*} v_i \;-\; K \cdot \sum_{i \in P^*} v'_i
    \;<\;
    \sum_{i \in P^*} K
    \;=\;
    |P^*|\cdot K
    \;\le\;
    nK
    \;=\;
    \epsilon\cdot V_{\max}^{\text{feas}}.
    \]
    Since there exists a feasible solution that takes one item achieving $V_{\max}^{\text{feas}}$, we have $OPT\ge V_{\max}^{\text{feas}}$ and therefore $nK\le \epsilon\cdot OPT$.

    Let $\widehat{P}$ be the set returned by the scaled DP (it maximizes $\sum_{i\in P} v'_i$ among feasible sets). Then
    \[
    \sum_{i\in \widehat{P}} v'_i \;\ge\; \sum_{i\in P^*} v'_i
    \;\;\Longrightarrow\;\;
    K\sum_{i\in \widehat{P}} v'_i \;\ge\; K\sum_{i\in P^*} v'_i.
    \]
    Using $\sum_{i\in P} v_i \ge K\sum_{i\in P} v'_i$ for every set $P$, we get
    \[
    \sum_{i\in \widehat{P}} v_i
    \;\ge\;
    K\sum_{i\in \widehat{P}} v'_i
    \;\ge\;
    K\sum_{i\in P^*} v'_i
    \;\ge\;
    \sum_{i\in P^*} v_i \;-\; \epsilon\cdot OPT
    \;=\;
    (1-\epsilon)\cdot OPT.
    \]
    Thus the algorithm is a $(1-\epsilon)$-approximation. (If desired, one can recover the actual set $\widehat{P}$ by storing predecessor pointers for each kept pair.)
\end{proof}
