\section*{Problem 5: My own question}
Propose a question related to the material covered in the course and
include a complete and formal solution. The level is expected to be similar
to the level of other questions in this assignment. Extremely well written and
elegant question may receive a bonus.

\subsection*{The question - The taxi problem}
Let $G = (V, E)$ be a complete graph. Let $t: V \times V \to \mathbb{R}_+$ be a metric representing the travel time between any two vertices. We are given an integer $B \geq 1$ representing the number of available taxis, and a set of $n$ passengers. Each passenger $i$ is defined by a tuple $(s_i, d_i) \in V \times V$, where $s_i$ is their source vertex and $d_i$ is their destination vertex.

All $B$ taxis have infinite passenger capacity, travel at a constant speed of 1, and begin their routes at time $0$ from any chosen starting vertices in $V$. Once a passenger is picked up by a taxi at their source $s_i$, they must remain in that same taxi until they are dropped off at their destination $d_i$ (i.e., transfers between taxis are not allowed).

A valid schedule defines the starting vertex and the sequence of visited vertices for each taxi over time such that every passenger $i$ is successfully transported from $s_i$ to $d_i$. Let $C_i$ denote the completion time (arrival time) of passenger $i$ at their destination $d_i$.

find valid schedules for all $B$ taxis that minimize the maximum completion time among all passengers: $\max_{i} C_i$.

Show that this problem is NP-hard, and show a $k$-approximation algorithm for the problem for $B=1$.

\subsection*{NP-hardness}

\begin{theorem}
The taxi problem is NP-hard, even for a single taxi ($B = 1$).
\end{theorem}

\begin{proof}
We show that the taxi problem is NP-hard even for a single taxi ($B = 1$), by reduction from the \textbf{Hamiltonian Path Problem}.

\medskip
\noindent\textbf{Reduction.}
Given an instance $H = (V_H, E_H)$ of Hamiltonian Path, we construct a taxi instance as follows.

\begin{itemize}
    \item \textbf{Vertices.} Let $V = V_H \cup \{d^*\}$, where $d^*$ is a fresh depot vertex not in $V_H$.

    \item \textbf{Metric.} Define the travel time $t$ as the shortest-path distance in the complete graph where:
    \begin{itemize}
        \item For $u, v \in V_H$: if $(u,v) \in E_H$, set $w(u,v) = 1$; otherwise $w(u,v) = 2$.
        \item For $v \in V_H$: set $w(v, d^*) = m^2$.
    \end{itemize}
    Then define $t(u,v)$ as the shortest-path distance under $w$. Since shortest-path distances satisfy the triangle inequality, $t$ is a valid metric, and it is computable in polynomial time.

    Note that for $u,v \in V_H$: $t(u,v) = 1$ if $(u,v) \in E_H$, and $t(u,v) = 2$ otherwise (the direct weight of $2$ is at most any two-hop path of cost $\geq 2$).

    \item \textbf{Taxis.} $B = 1$.

    \item \textbf{Passengers.} For each vertex $v_i \in V_H$, create a passenger with source $s_i = v_i$ and destination $d_i = d^*$.

    \item \textbf{Threshold.} $T = (m - 1) + m^2$.
\end{itemize}

\noindent\textbf{Correctness.}

In any valid schedule, the single taxi must visit every source $v_i \in V_H$ to pick up all $m$ passengers, and then reach $d^*$ to drop them off. Since the taxi has infinite capacity, the optimal strategy is to first tour all of $V_H$ (picking up everyone), and then travel to $d^*$ (dropping everyone off). The makespan equals
\[
    C_{\max} = (\text{length of path through all vertices of } V_H) + t(v_{\text{last}}, d^*),
\]
where $v_{\text{last}}$ is the last vertex of $V_H$ visited. Since $t(v, d^*) = m^2$ for every $v \in V_H$, this simplifies to
\[
    C_{\max} = (\text{length of path through all vertices of } V_H) + m^2.
\]

\begin{itemize}
    \item[$(\Rightarrow)$] If $H$ has a Hamiltonian path $v_{\pi(1)}, v_{\pi(2)}, \ldots, v_{\pi(m)}$, the taxi starts at $v_{\pi(1)}$ and follows this path. Each edge has $t$-cost $1$ (since consecutive vertices are adjacent in $H$), so the path has total length $m - 1$. The end time is $(m-1) + m^2 = T$.

    \item[$(\Leftarrow)$] Suppose there is a valid schedule with $\leq T = (m-1) + m^2$. The taxi visits all $m$ vertices of $V_H$ in some order, traversing $m - 1$ edges, then travels to $d^*$ at cost $m^2$. The total path length through $V_H$ must be $\leq m - 1$. Since each step between consecutive visited vertices costs at least $1$ (as $t(u,v) \geq 1$ for all $u \neq v$ in $V_H$), and there are exactly $m - 1$ steps, each step must cost exactly $1$. By construction, $t(u,v) = 1$ if and only if $(u,v) \in E_H$. Therefore the sequence of visited vertices forms a Hamiltonian path in $H$.
\end{itemize}

The reduction runs in polynomial time (constructing the metric takes $O(m^2)$ time). Since the Hamiltonian Path Problem is NP-complete, the taxi problem (even with $B=1$) is NP-hard.
\end{proof}

\subsection*{K approximation algorithm for the taxi problem}
\textbf{Algorithm.} the idea for greedy algorithm for the general case is the following. the time for a taxi with the passengers list $\{a_1, \ldots, a_m\}$ is the TSP approximation of the path for the sources, then the shortest path from the last source to the first destination, then the TSP approximation of the path for the destinations. at each iteration, we assign the next passenger such that the increase in the completion time of all the taxis is minimal. we repeat until all passengers are assigned.

here is the algorithm for the single taxi case ($B=1$):
\begin{algorithm}
\caption{4-Approximation for Single Taxi ($B=1$)}
\begin{algorithmic}[1]
\Procedure{SingleTaxiSchedule}{$G, t, \{(s_1, d_1), \ldots, (s_n, d_n)\}$}
    \State $S \gets \{s_1, \ldots, s_n\}$
    \State $D \gets \{d_1, \ldots, d_n\}$

    \State \textbf{Phase 1: Collect all passengers}
    \State $Path_S \gets \text{ChristofidesPath}(S, t)$
    \State $s_{last} \gets \text{last vertex in } Path_S$

    \State \textbf{Phase 2: Transition}
    \State $d_{first} \gets \text{Find } d_i \in D \text{ that minimizes } t(s_{last}, d_i)$

    \State \textbf{Phase 3: Deliver all passengers}
    \State $Path_D \gets \text{ChristofidesPath}(D, t, \text{start}=d_{first})$

    \State \Return $Path_S \oplus Path_D$ \Comment{Concatenate the two paths}
\EndProcedure
\end{algorithmic}
\end{algorithm}


\textbf{Complexity.} The algorithm runs in polynomial time, since Christofides' algorithm runs in polynomial time, and finding the minimum $t(s_{last}, d_i)$ takes $O(n)$ time.


\textbf{Approximation Analysis.}

\begin{claim}
The greedy algorithm is a 4-approximation algorithm for the single taxi problem ($B=1$).
\end{claim}

\begin{proof}
Let $OPT$ be the optimal maximum completion time (makespan) for a single taxi to deliver all passengers. Let $S = \{s_1, \dots, s_n\}$ be the set of all source vertices and $D = \{d_1, \dots, d_n\}$ be the set of all destination vertices.

\noindent\textbf{1. Lower Bounds on OPT:}\\
Any valid optimal schedule must visit all vertices in $S$ to pick up the passengers, and it must visit all vertices in $D$ to drop them off. Because the travel times $t$ form a metric (satisfying the triangle inequality), shortcutting any intermediate vertices can only decrease the total path length. Therefore, we can establish the following lower bounds:
\begin{itemize}
\item \textbf{Lower Bounds 1 \& 2 (Sources and Destinations):} The optimal schedule contains a path that visits all vertices in $S$ and all vertices in $D$. Thus, the shortest paths covering all sources and destinations, denoted $OPT_{path}(S)$ and $OPT_{path}(D)$, are both lower bounds on $OPT$:
$$OPT \geq OPT_{path}(S) \quad \text{and} \quad OPT \geq OPT_{path}(D)$$
    \item \textbf{Lower Bound 3 (Max Edge):} The optimal schedule must visit every node in $S \cup D$. By the triangle inequality, the distance between \emph{any} two nodes in $S \cup D$ is at most the length of the optimal path. We have: $OPT \geq t(s_{last}, d_{first})$
\end{itemize}

\noindent\textbf{2. Upper Bound on the Algorithm's Cost:}\\
The proposed algorithm constructs a schedule in three sequential phases:
\begin{enumerate}
    \item A TSP path covering $S$.
    \item A single transit edge from $s_{last}$ to $d_{first}$.
    \item A TSP path covering $D$.
\end{enumerate}

Let $ALG$ be the total completion time of this algorithm. The cost is the sum of the three phases:
$$ALG = Cost(Path_S) + t(s_{last}, d_{first}) + Cost(Path_D)$$

We are assuming the use of Christofides' algorithm adapted for the Shortest Hamiltonian Path problem with free endpoints. This algorithm guarantees a $1.5$-approximation for the optimal path. Therefore:
\begin{itemize}
    \item $Cost(Path_S) \leq 1.5 \cdot OPT_{path}(S)$
    \item $Cost(Path_D) \leq 1.5 \cdot OPT_{path}(D)$
\end{itemize}

\noindent\textbf{3. Combining Bounds:}\\
Substitute the approximations and the lower bounds into the algorithm's cost equation:
$$ALG \leq 1.5 \cdot OPT_{path}(S) + t(s_{last}, d_{first}) + 1.5 \cdot OPT_{path}(D)$$

Using our lower bounds ($OPT_{path}(S) \leq OPT$, $t(s_{last}, d_{first}) \leq OPT$, and $OPT_{path}(D) \leq OPT$):
$$ALG \leq 1.5 \cdot OPT + 1 \cdot OPT + 1.5 \cdot OPT$$
$$ALG \leq 4 \cdot OPT$$

Thus, the proposed greedy algorithm is a \textbf{4-approximation algorithm} for the single taxi problem.
\end{proof}
